---
title: "US_WS_Recoding"
author: "Jerry"
date: "September 24, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('xlsx')
library("readxl")
library('tidyverse')
library('scales')
library('lazyeval')
```


#########################
#### US Data loading ####
#########################
```{r 1. Dataset}
#### US dataset ####
dataset_name = 'dat_well0_2020_1215.csv'
setwd("C:/Users/WeiKuang/Desktop/Stanford Work/WELL/Data/Data_US/")
DF_US = read.csv(dataset_name, stringsAsFactors = FALSE)
dim(DF_US)
names(DF_US) # Old well score is already here!
```




##################
#### Recoding ####
##################
```{r Site}
DF_US$Site = 'US'
DF_US$Site  = factor(DF_US$Site)
```

```{r Covariates -  Demographic Variable}

#### Age-1####
X = DF_US$portal_age %>% cut( breaks= c(18, 30, 50, 60, 70, 120),right = F) %>% unlist() %>% as.vector()
X %>% table(useNA='always')
X = tidyr::replace_na(X, 'Missing')
X_fct = factor(X, labels = c('18-29','30-49','50-59','60-69','70+'))
X_fct %>% table(useNA = 'always')
DF_US$age_new_1 = X_fct

#### Age-2 ####
X = DF_US$portal_age %>% cut( breaks= c(0, 50, 60, 70, 80, 120),right = F) %>% unlist() %>% as.vector()
X %>% table(useNA='always')
X = tidyr::replace_na(X, 'Missing')
X_fct = factor(X, labels = c('<50','50-59','60-69','70-79','80+'))
X_fct %>% table(useNA = 'always')
DF_US$age_new_2 = X_fct


#### Gender/Sex-1 ####
my_var = 'core_gender'
table(DF_US[[my_var]], useNA = 'always')
Y = DF_US[[my_var]]
X = DF_US[[my_var]]
X[Y==5] <- '1, Male'
X[Y==4] <- '2, Female'
X[Y %in% c(1,2,3,6)] <- NA  #'3, Other'  -> NA
X = tidyr::replace_na(X, 'Missing')
X_fct = factor(X); X_fct %>% table(useNA = 'always')
DF_US$gender_new = X_fct


#### Gender/Sex-2 ####
my_var = 'core_gender'
table(DF_US[[my_var]], useNA = 'always')
Y = DF_US[[my_var]]
X = DF_US[[my_var]]
X[Y==5] <- '1, Male'
X[Y==4] <- '2, Female'
X[Y %in% c(1,2,3,6)] <- '3, Other'  
X = tidyr::replace_na(X, 'Missing')
X_fct = factor(X); X_fct %>% table(useNA = 'always')
DF_US$gender_new_MFO = X_fct

#### Education ####
#### function - education ####
function_education = function(data_input, timepoint_suffix ='_well0'){
  
  var1 = 'core_education_us_level'
  var2 = 'core_education_nonus_level'
  var1.time = paste0(var1, timepoint_suffix)
  var2.time = paste0(var2, timepoint_suffix)
  
  DF = data_input
  DF$edu1<-DF[[var1.time]]
  DF$edu2<-DF[[var2.time]]
  
  #### dv_edu_3cat ####
  vector= rep(NA, nrow(DF))
  vector[DF$edu1==1|DF$edu1==2|DF$edu1==3|DF$edu1==4|DF$edu1==5|DF$edu1==6|DF$edu1==7|
              DF$edu2==1|DF$edu2==2|DF$edu2==3|DF$edu2==4|DF$edu2==5]<-1 #High school and under
  
  vector[DF$edu1==8|DF$edu1==9|DF$edu1==10|
              DF$edu2==6|DF$edu2==7|DF$edu2==8]<-2 #College
  
  vector[DF$edu1==11|DF$edu1==12|DF$edu1==13|DF$edu2==9]<-3 #Master's and above
  
  vector_fct <- factor(vector,
                      levels = c(1,2,3))
                      # labels = c("1, High school and under","2, College","3, Master's and above"))
  
  DF$dv_edu_3cat = vector_fct
  
  
  
  
  #### dv_edu_4cat ####
  vector = rep(NA, nrow(DF))
  vector[DF$edu1%in%c(1,2,3,4,5)|DF$edu2%in%c(1,2,3,4)]<-1 #Primary school and under
  vector[DF$edu1%in%c(6,7)|DF$edu2%in%5]<-1 #High School/GED or completed Secondary school
  vector[DF$edu1%in%c(8,9)|DF$edu2%in%c(6,7)]<-2 #Some college, no degree or Associate/Technical/Vocational education
  vector[DF$edu1%in%10|DF$edu2%in%8]<-3 #Bachelors/university level
  vector[DF$edu1%in%c(11,12,13)|DF$edu2%in%9]<-4 #Post-Graduate/Professional
  vector_fct <- factor(vector,
                   levels = c(1,2,3,4))
                   #labels = c("1, High school and under",
                   #           "2, Some college, no degree or Associate/Technical/Vocational education",
                   #           "3, Bachelors/university level",
                   #           "4, Post-Graduate/Professional"))
  DF$dv_edu_4cat = vector_fct
  
  
  #### dv_edu_5cat ####
  vector = rep(NA, nrow(DF))
  vector [DF$edu1%in%c(1,2,3,4,5)|DF$edu2%in%c(1,2,3,4)]<-1 #Primary school and under
  vector [DF$edu1%in%c(6,7)|DF$edu2%in%5]<-2 #High School/GED or completed Secondary school
  vector [DF$edu1%in%c(8,9)|DF$edu2%in%c(6,7)]<-3 #Some college, no degree or Associate/Technical/Vocational education
  vector [DF$edu1%in%10|DF$edu2%in%8]<-4 #Bachelors/university level
  vector [DF$edu1%in%c(11,12,13)|DF$edu2%in%9]<-5 #Post-Graduate/Professional
  vector_fct  <- factor(vector,
                       levels = c(1,2,3,4,5))
                       #labels = c("1, Some high school and under",
                       #            "2, High School graduate/GED or completed Secondary school",
                       #            "3, Some college, no degree or Associate/Technical/Vocational education",
                       #            "4, Bachelors/university level",
                       #            "5, Post-Graduate/Professional"))
  
  DF$dv_edu_5cat = vector_fct
  
  #### output
  df_out = DF %>% select('id','dv_edu_3cat','dv_edu_4cat','dv_edu_5cat')
  #### Rename the column
  colnames(df_out)[2:4] = paste0(  colnames(df_out)[2:4], timepoint_suffix)
  return(df_out)
}



#### education - recoding ####
df = function_education(data_input = DF_US, timepoint_suffix ='')
X = df$dv_edu_4cat 
X = as.character(X)
X = tidyr::replace_na(X, 'Missing')
X %>% table(useNA = 'always')

X_fct <- factor(X,
                levels = c('1','2','3','4', 'Missing'),
                labels = c("1, High school and under",
                           "2, Some college or Associate/Vocational Education.",
                           "3, Bachelors/university level",
                           "4, Post-Graduate/Professional",
                           "Missing")
                ) 

X_fct %>% table(useNA = 'always')
DF_US$edu_new = X_fct


#### Income -1  ####
DF_US$core_income %>% summary()

DF_US$core_income_NEW_3cat =NA
DF_US$core_income_NEW_3cat  [DF_US$core_income %in% c(0:6)]  = "< 100K"
DF_US$core_income_NEW_3cat  [DF_US$core_income %in% c(7,8)]  = "100-249K"
DF_US$core_income_NEW_3cat  [DF_US$core_income %in% c(9,10,11)] = "250K+"

table(DF_US$core_income_NEW_3cat,useNA = 'always')
DF_US$core_income_NEW_3cat = factor(DF_US$core_income_NEW_3cat, levels= c( "< 100K", "100-249K", "250K+"))
DF_US$core_income_NEW_3cat = fct_explicit_na(DF_US$core_income_NEW_3cat, na_level = 'Missing')
DF_US$core_income_NEW_3cat %>% table(useNA = 'always')


#### Income -2  ####
DF_US$core_income %>% summary()

X = rep(NA, nrow(DF_US)) 
Y = DF_US$core_income 

X[Y %in% c(0:6)]  = "< 100K"
X[Y %in% c(7,8)]  = "100-249K"
X[Y %in% c(9)]  = "250-499K"
X[Y %in% c(10,11)]  = "500+K"
X %>% table()
X_fct = factor(X)
DF_US[['core_income_NEW_4cat']]  = X


#### Employment/Job/Occupation ####
  # 1, Working now 
  # 2, Temporarily laid off
  # 3, Sick leave or maternity leave 
  # 4, Unemployed, looking for work 
  # 5, Retired 
  # 6, Disabled 
  # 7, Home maker 
  # 8, Student 
  # 9, Other
my_var = 'core_work_status'
table(DF_US[[my_var]], useNA = 'always')
Y = DF_US[[my_var]]
X = DF_US[[my_var]]
X[Y==1] <- "Working now"
X[Y==2] <- "Temporarily laid off"
X[Y==3] <- "Sick leave or maternity leave"
X[Y==4] <- "Unemployed, looking for work"
X[Y==5] <- "Retired"
X[Y==6] <- "Disabled"
X[Y==7] <- "Home maker"
X[Y==8] <- "Student"
X[Y==9] <- "Other"
X_fct = factor(X)
X_fct = fct_explicit_na(X_fct,na_level = 'Missing')
X_fct %>% table(useNA='always')
DF_US$core_work_status_new = X_fct
```

```{r race - Jerry's checking}
dim(DF_US)

#### Pure CN and mixed ####
DF_US %>% filter(core_race___9==1) %>% nrow()

# Pure CN
DF_US %>% dplyr::filter(
  core_race___1==0,
  core_race___2==0,
  core_race___3==0,
  core_race___4==0,
  core_race___5==0,
  core_race___6==0,
  core_race___7==0,
  core_race___8==0,
  core_race___9==1,
  core_race___10==0,
  core_race___11==0,
  core_race___12==0,
  core_race___13==0,
  core_race___14==0,
  core_race___15==0,
  core_race___16==0) %>% 
  nrow()


#### Pure CN and mixed ####
DF_US %>% filter(core_race___15==1) %>% nrow()

# Pure TW
DF_US %>% dplyr::filter(
  core_race___1==0,
  core_race___2==0,
  core_race___3==0,
  core_race___4==0,
  core_race___5==0,
  core_race___6==0,
  core_race___7==0,
  core_race___8==0,
  core_race___9==0,
  core_race___10==0,
  core_race___11==0,
  core_race___12==0,
  core_race___13==0,
  core_race___14==0,
  core_race___15==1,
  core_race___16==0) %>% 
  nrow()


#### CN & TW Only ####
DF_US %>% dplyr::filter(
  core_race___1==0,
  core_race___2==0,
  core_race___3==0,
  core_race___4==0,
  core_race___5==0,
  core_race___6==0,
  core_race___7==0,
  core_race___8==0,
  core_race___9==1,
  core_race___10==0,
  core_race___11==0,
  core_race___12==0,
  core_race___13==0,
  core_race___14==0,
  core_race___15==1,
  core_race___16==0) %>%  dplyr::select(starts_with("core_race___") )

```
```{r race - pre-encoding}
#### Action: Convert the "1" into a representative text
#### For example: If core_race =1, then I want "1,White" as the value.

#### Why to do so? Race/Ethinicity allows people to check multiple choice at the same time.
#### Thus, I need to capture "the combination of every races which were selected by a participant"



#### Step-1: Define my wanted text/label ####
my_lables = c(
'1,White',  
'2,AfricanAmerican', 
'3,NativeAmerican',   
'4,NativeAlaska',  
'5,NativeHawaiian',   
'6,Guamanian',   
'7,Samoan',    
'8,OtherPacificIslander',   
'9,Chinese',   
'10,Japanese',   
'11,Korean',  
'12,Vietnamese',  
'13,OtherAsian',   
'14,OtherRace',
'15,Taiwanese',   
'16,SouthAsian'
) 



#### Step-2: Create a place holder "df" to store the recorded race/ethnicity
checkbox_var = paste0('core_race___', c(1:16))
df = DF_US %>% dplyr::select( c('id', checkbox_var) ) 

for (i in 1:length(checkbox_var) ) {
  print(i)
  print(my_lables[i])
  new_var_name = paste0(checkbox_var[i],'_new')
  
  # Create place holder
  df[[new_var_name]] = rep(NA, nrow(df))
  df[[new_var_name]][df[[checkbox_var[i]]]==1] <- my_lables[i]
}
    #### check - This df has the original core_race[0-16] and recoded core_race[0-16]_new
    head(df)

    
#### Step-3: I only need the recoded 'core_race___[0-9]*_new'
select= colnames(df) %>% grepl(x =., pattern = 'core_race___[0-9]*_new')
my_col = colnames(df)[select]
df[, my_col] %>% head()
    
#### Great! In the next section, I will combine all the text row by row ####
```

```{r race - recoding-1}
#### Race is a check-box variable which allows a person to give multiple responses
#### That's why we need to check all combination

#### data manipulation 
string_vector = df %>% 
  dplyr::select(my_col) %>%
  apply(X = .,MARGIN = 1,FUN =function(x){ paste(x,collapse = '_')}) %>% # All combination
  str_replace_all(string =.,pattern = '[0-9]*,|NA', replacement = '')%>%
  str_replace_all(string =.,pattern = '_+', replacement = ' ')%>%
  str_replace_all(string =.,pattern = '^ | $', replacement = '')

string_vector [string_vector=='']<-NA
    #### check ####
    string_vector %>% table(useNA = 'always')



#### Assign back to the WELL US master dataset ####
# Note: I did not remove any rows or change the position, so df and DF_US have the same row.
# Thus, id or join were not used to merge the data.
    #### check ####
    nrow(DF_US)
    nrow(df)

df$core_race_new = string_vector
DF_US$core_race_new = string_vector

#### Take care of hispanic #####
# Note: core_race did not ask Hispanic, but it is core_ethnic asking Hispanic.
# Thus, [DF_US$core_ethnic ==1] ='Hispanic'  is the only place I can confirm  "Hispanic"
# Not perfect, but this is what I can do with this survey data. 

DF_US$core_race_new [DF_US$core_ethnic ==1] ='Hispanic'
DF_US$core_race_new   %>% table(useNA = 'always')

### END of Ethnicity/Race coding ####
```

```{r race - Jerry's checking for pure and mixed race/ethnicity }
# pure race -> which has no ' '
# n= ~ 5600
select_pure_race = ! grepl(x=df$core_race_new , pattern= ' ')
sum(select_pure_race)
df$core_race_new [select_pure_race]  %>% table(useNA = 'always')

# mixed (n=~600)
select_mixed_race =  grepl(x=df$core_race_new , pattern= ' ')
sum(select_mixed_race)
df$core_race_new [select_mixed_race]  %>% table(useNA = 'always')
```



#### Check Ethnic groups in Asain race ####
```{r Ethinicity  in Asian-WRONG}
DF_US$race_NEW.temp %>% table(useNA = 'always')
setwd('C:/Users/WeiKuang/Desktop/Stanford Work/WELL/Project/WELL_Score/Result')
DF_US %>% count(race_NEW,.drop = FALSE)  %>% write.xlsx('US_race.xlsx')


DF_US%>% 
  filter( race_NEW.temp %in%c (13 ) )%>%    # 13,Other Asian
  count(core_asian_other)

DF_US%>% 
  filter( race_NEW.temp %in%c (16 ) )%>%  # 16,South Asian  
  count(core_asian_other)

#### recoding ####
X = rep(NA, nrow(DF_US))
Y1 = DF_US$race_NEW.temp
Y2 = DF_US$core_asian_other

X[Y1 == 9] <- '9,Chinese'
X[Y1 == 10] <- '10,Japanese'
X[Y1 == 11] <- '11,Korean'
X[Y1 == 12] <- '12,Vietnamese'
X[Y1 == 15] <- '15,Taiwanese'

#### '13,Other Asian' and '16,South Asian'
X[Y1 == 13 & (Y2 %in% 'Cambodian')] <- 'Cambodian'
X[Y1 == 13 & (Y2 %in% 'Taiwanese')] <- '15,Taiwanese'

my_Indian = c('India','indian', 'Indian')
X[Y1 == 13 & (Y2 %in% my_Indian )] <- 'Indian'

my_Pilipino = c('Filipina','filipino', 'Filipino', 'Filipinx','Fillipino','Filpino','FILPINO', 'Pilipino')
X[Y1 == 13 & (Y2 %in% my_Pilipino)] <- 'Pilipino/Filipino'

#### clean up - these people say they are lation
X [DF_US$core_ethnic ==1] = NA

X %>% table()

#### Assign - 1
DF_US[['Ethinicity_Asian_new']] = X
df = DF_US %>% count(Ethinicity_Asian_new)
df

#### Final recoding 
X = as.character(DF_US$race_NEW)
Y3 = DF_US$Ethinicity_Asian_new

X [ Y3 == 'Indian'] = 'Indian'
X [ Y3 == 'Pilipino/Filipino'] = 'Pilipino/Filipino'
X %>% table()

DF_US[['race_NEW_more_asian']] = X

setwd('C:/Users/WeiKuang/Desktop/Stanford Work/WELL/Project/WELL_Score/Result')
DF_US %>% count(race_NEW_more_asian,.drop = FALSE)  %>% write.xlsx('US_race.xlsx')
```





#### Other Covariates #####
```{r live alone at WELL0}
#### Live along ####
var= 'core_pplno_household'
DF_US[[var]] %>% table(useNA = 'always')
X=DF_US[[var]] 
Y=DF_US[[var]] 
X[Y==1] <- 'Yes'
X[Y!=1] <- 'No'
X[is.na(Y)] <- NA
X_fct = factor(X)
X_fct %>% table(useNA = 'always')
DF_US[['core_pplno_household_new']] = X_fct
```

```{r Mobility}
my_var = 'core_nature_min'
table(DF_US[[my_var]], useNA = 'always')
Y = DF_US[[my_var]]
X = DF_US[[my_var]]
X[Y==1] <- '1, Never or Alost never'
X[Y==2] <- '1, Never or Alost never'
X[Y==3] <- '2, Sometimes'
X[Y==4] <- '3, Fairly often'
X[Y==5] <- '4, Very often'
X_fct = factor(X)
X_fct = fct_explicit_na(X_fct, na_level = 'Missing')
X_fct %>% table(useNA = 'always')
DF_US$core_nature_min_new = X_fct
```

```{r Disease: diabetes, hypertension, a morbidity index }

#### Diabetes ####
  my_var = 'core_diabetes'
  DF_US[, my_var] %>% table(useNA = 'always')
  Y = DF_US[[my_var]]
  X = DF_US[[my_var]]
  X[Y==0] <- 'No'
  X[Y==1] <- 'Yes'
  X[Y==2] <- NA
  X = tidyr::replace_na(X, 'Missing')
  X_fct = factor(X, levels = c('No','Yes','Missing'))
  X_fct %>% table(useNA = 'always')
  DF_US$core_diabetes_new = X_fct

 
#### hypertension ####
  my_var = 'core_hypertension'
  DF_US[, my_var] %>% table(useNA = 'always')
  Y = DF_US[[my_var]]
  X = DF_US[[my_var]]
  X[Y==0] <- 'No'
  X[Y==1] <- 'Yes'
  X[Y==2] <- NA
  X = tidyr::replace_na(X, 'Missing')
  X_fct = factor(X, levels = c('No','Yes','Missing'))
  X_fct %>% table(useNA = 'always')
  DF_US$core_hypertension_new = X_fct

  
#### cancer - History of cancer ####
  my_var = 'core_cancer'
  DF_US[, my_var] %>% table(useNA = 'always')
  Y = DF_US[[my_var]]
  X = DF_US[[my_var]]
  X[Y==0] <- 'No'
  X[Y==1] <- 'Yes'
  X[Y==2] <- NA
  X = tidyr::replace_na(X, 'Missing')
  X_fct = factor(X, levels = c('No','Yes','Missing'))
  X_fct %>% table(useNA = 'always')
  DF_US$core_cancer_new = X_fct


#### Correction Comorbidity 04/15/2020 ####
function_comorbidity_index = function(data_input, 
                                      new_var_name = 'dv_comorbidity_index',
                                      timepoint_suffix='_well0'){
  
  Medical_Var = c(
    'core_hypertension','core_heartattack','core_heartdisease',
    'core_stroke','core_cancer','core_diabetes')
  
  Medical_Var = paste0(Medical_Var,timepoint_suffix)
  
  #### Corrected Comorbidity 04/15/2020 ####
  #### Convert disease status into numeric ( 1 = Yes,  0= any other values)
  DF_Dis = data_input %>% select(Medical_Var)
  DF_Dis [DF_Dis == 0| DF_Dis==2 | is.na(DF_Dis) ] =0
  
  #### Sum up to create the comorbidity index
  DF_Dis[[ new_var_name]] = rowSums(DF_Dis,na.rm = T)

  
  #### Assign 
  data_input[[ new_var_name]] = DF_Dis[[ new_var_name]]

  #### Output
  df_out = data_input %>% select('id',new_var_name )
  #### Rename
  colnames(df_out)[2] = paste0(new_var_name, timepoint_suffix)
  return(df_out)
}

#### Comorbidity  recoding ####
  df = function_comorbidity_index(data_input = DF_US,new_var_name = 'comorbidity_index', timepoint_suffix='')
  df$comorbidity_index %>% table(useNA = 'always')
  X = df$comorbidity_index 
  X = cut(X, breaks = c(0,1 ,2, 7), right = F)  
  X %>%  table(useNA = 'always' )
  X_fct = factor(X, levels= c('[0,1)','[1,2)','[2,7)'), labels = c('0', '1','2-6'))
  X_fct = fct_explicit_na(X_fct, na_level = 'Missing')
  X_fct %>% table(useNA = 'always')
  DF_US$comorbidity_index_new = X_fct
```

```{r Mental Condition}
#### Depression ####
  my_var = 'core_depression'
  DF_US[, my_var] %>% table(useNA = 'always')
  Y = DF_US[[my_var]]
  X = DF_US[[my_var]]
  X[Y==0] <- 'No'
  X[Y==1] <- 'Yes'
  X[Y==2] <- NA
  X = tidyr::replace_na(X, 'Missing')
  X_fct = factor(X, levels = c('No','Yes','Missing'))
  X_fct %>% table(useNA = 'always')
  DF_US$core_depression_new = X_fct
  
  
  
#### general mental health condition
X = rep(NA, nrow(DF_US))
Y1 = DF_US$core_depression
Y2 = DF_US$core_mental_health

X[ Y1 ==0 & Y2==0] <- '1. -- ' # -- None of any mental condition
X[ Y1 ==0 & Y2==1] <- '2. -+ ' # -+ Other mental health conditions
X[ Y1 ==1 & Y2==0] <- '3. +- ' # +- Clinical depression
X[ Y1 ==1 & Y2==1] <- '4. ++ ' # ++ Clinical depression + Other Mental health conditions


X_fct = factor(X)
X_fct %>% table(useNA = 'always')
DF_US$mental_condition_new = X_fct
DF_US$mental_condition_new %>% table(useNA = 'always')
```

```{r Life Style: smoking, drinking, and sleeping}
#### smoke ####
function_smoke= function(data_input, 
                         new_var_name='dv_smoke_current',
                         timepoint_suffix = '_well0') {
  library('dplyr')
  
  # Function annotations
  #### data_input must be US dataset
  #### Two variable are reuqied to derive this variable 
  #### var.old.1 = 'core_smoke_100'
  #### var.old.2 = 'core_smoke_freq' 
  #### derived responses: 0, Never | 1, Former| 2, Current

  dfus = data_input
  var.old.1 = paste0('core_smoke_100', timepoint_suffix)
  var.old.2 = paste0('core_smoke_freq', timepoint_suffix)
  checkpoint = ( c(var.old.1, var.old.1) %in% colnames(dfus) ) %>% sum()
  if(checkpoint <2){
    print('core_smoke_100 or core_smoke_freq  does not exist in the selected time point')
  }
  
  #### Ann's version for smoking status in US 
  vector = rep(NA, nrow(dfus))
  
  #### My Old program 
  #vector<-ifelse(dfus$core_smoke_100 %in%c(0), 0, #Never
  #                              ifelse(dfus$core_smoke_100%in%1 & dfus$core_smoke_freq%in%1, 1, #Former
  #                                     ifelse(dfus$core_smoke_freq%in% c(2,3), 2, #Current 
  #                                            NA)))

  vector<-ifelse(dfus[[var.old.1]] %in%c(0), 0, #Never
                ifelse(dfus[[var.old.1]]%in%1 & dfus[[var.old.2]] %in%1, 1, #Former
                       ifelse(dfus[[var.old.2]]%in% c(2,3), 2, #Current 
                                                            NA)))
  
  
  #### Label
  # vector_fct = factor(vector, levels=c(0,1,2), labels = c('Never','Former','Current'))
  vector_fct = factor(vector, levels=c(0,1,2)) # 0, Never | 1, Former| 2, Current
  dfus[[new_var_name]] = vector_fct

  #### print the table for quick checking
  print(table(dfus[[new_var_name]], useNA = 'always'))
  
  #### output
  df_out = dfus %>% dplyr::select('id',new_var_name)
  #### Add time suffix to the new-derived varaible
  colnames(df_out)[2] = paste0(new_var_name, timepoint_suffix)
  return(df_out)
}

#### smoke recoding ####
  df = function_smoke(data_input = DF_US, timepoint_suffix = '') # note: output is factor! 
  X = df$dv_smoke_current
  X_fct = factor(X, levels = c('0','1','2'), labels = c('0, Never', '1, Former','2, Current'))
  X_fct %>% table(useNA = 'always')
  X_fct = fct_explicit_na(X_fct, na_level = 'Missing')
  X_fct %>% table(useNA = 'always')
  DF_US$smoke_new = X_fct
  
#### drinking ####
function_alcohol = function(data_input, 
                            new_var_name='dv_alcohol_current', 
                            timepoint_suffix = '_t0') {
  
  # Function annotations
  #### data_input must be US dataset
  #### 'core_alcohol_day' is the variable
  #### Output is a data frame with the derived drinking status
  #### Derived: 0, Non-drinkers | 1, Current
  
  var.old = paste0('core_alcohol_day', timepoint_suffix)
  
  dfus = data_input
  vector = rep(NA, nrow(dfus))
  vector<-ifelse(dfus[[var.old]] %in%0, 0, # Non-drinkers
                              ifelse(dfus[[var.old]]%in%c(1:30), 1, #Current
                                     NA))

  #### label
  # vector_derived = factor(vector, levels=c(0,1), labels = c('Non-drinkers','Current'))
  vector_derived = factor(vector, levels=c(0,1)) # 0, Non-drinkers | 1, Current
  dfus[[new_var_name]] = vector_derived
  
  #### print the table for quick checking
  print(table( dfus[[new_var_name]], useNA = 'always'))
  
  #### output
  df_out = dfus %>% dplyr::select('id', new_var_name )
  
  #### Add time suffix to the new-derived varaible
  colnames(df_out)[2] = paste0(new_var_name, timepoint_suffix)
  return(df_out)
}
  
#### smoke recoding ####
  df = function_alcohol(data_input = DF_US, timepoint_suffix = '') # note: output is factor! 
  X = df$dv_alcohol_current
  X_fct = factor(X, levels = c('0','1'), labels = c('0, Non-drinkers', '1, Current'))
  X_fct %>% table(useNA = 'always')
  X_fct = fct_explicit_na(X_fct, na_level = 'Missing')
  X_fct %>% table(useNA = 'always')
  DF_US$drinking_new = X_fct  
  
#### sleep ( psqi is not available )####
  my_var = 'core_sleep_quality'
  DF_US[, my_var] %>% table(useNA = 'always')
  Y = DF_US[[my_var]]
  X = DF_US[[my_var]]
  X[Y==1] <- '1, Very bad'
  X[Y==2] <- '2, Fairly bad '
  X[Y==3] <- '3, Fairly good '
  X[Y==4] <- '4, Very good'
  X_fct = factor(X)
  X_fct = fct_explicit_na(X_fct, na_level = 'Missing')
  X_fct %>% table(useNA = 'always')
  DF_US$core_sleep_quality_new = X_fct
```

```{r Physical Activity}
# How often did you do something outside for a period of time lasting more than 10 minutes?
DF_US$core_nature_min %>% table(useNA = 'always')

X=rep(NA, nrow(DF_US))
Y=DF_US$core_nature_min

X[Y %in% c(1,2)] ='1. Never and Almost never'
X[Y %in% c(3)]   ='2. Sometimes'
X[Y %in% c(4)]   ='3. Fairly often'
X[Y %in% c(5)]   ='4. Very often'
X %>% table(useNA = 'always')
X_fct = factor(X)
DF_US['core_nature_min_new'] = X_fct
```



#### country of residence & Zipcode ####
```{r archive - zipcode preparation}
# Jerry: Convert 'zip_code_database.csv' into "zip_code_database_JerryCleaned.rds"

# setwd("C:/Users/WeiKuang/Desktop/Stanford Work/WELL/Project/COVID19/Data/")
# list.files()

# zipcode_US =  read.csv("zip_code_database.csv")
# zipcode_US$zip = as.character(zipcode_US$zip)
# 
# zipcode_US$zip_cleaned = zipcode_US$zip
# select.1 = nchar(zipcode_US$zip)==3
# zipcode_US$zip_cleaned [select.1] = paste0('00',zipcode_US$zip [select.1])
# select.2 = nchar(zipcode_US$zip)==4
# zipcode_US$zip_cleaned [select.2] = paste0('0',zipcode_US$zip [select.2])
# 
# 
# #### replace
# zipcode_US$zip = as.character(zipcode_US$zip_cleaned )
# setwd("C:/Users/WeiKuang/Desktop/Stanford Work/WELL/Project/COVID19/Data/")
# zipcode_US$zip = as.character(zipcode_US)
# zipcode_US %>% select(zip, primary_city, state, county, country) %>% 


#### Output #####
# setwd("C:/Users/WeiKuang/Desktop/Stanford Work/WELL/Project/COVID19/Data/")
# list.files()
# zipcode_US = readRDS('zip_code_database_JerryCleaned.rds')
# str(zipcode_US)
```

```{r merge to US master file}
####  US zipcode
setwd("C:/Users/WeiKuang/Desktop/Stanford Work/WELL/Project/COVID19/Data/")
zipcode_US =  readRDS('zip_code_database_JerryCleaned.rds')
str(zipcode_US) # This dataset has zip, city, state, county, and country (US only)

#### merge
DF_US = left_join(x=DF_US , y=zipcode_US , by=c('portal_zip'='zip'))

# Check
# DF_US %>% select('county') %>% is.na() %>% sum()
# DF_US %>% filter(is.na(county))  %>% select(portal_zip, county)
```



#### WELL score Derivation ####
```{r}
#### File Path ####
setwd("C:/Users/WeiKuang/Desktop/Stanford Work/WELL/Project/David_FactorAnalysis/Code/1_US_CFA/WELL_Score_function/WELL_Score_Function/version_09.23.20/")
source('0923.function_well_score_US_model7_version2.R')


#### Run ####
# df_version2 = f_score_US_model_7_v2(data.arg = DF_US)
# summary(df_version2)
#### Merger ####
# DF_US = left_join(x=DF_US,y=df_version2, by='id')
```


#### Save the recoded dataset ####
```{r}
getwd()

#### Project - WELL CFA ####
# my_depo = "C:/Users/WeiKuang/Desktop/Stanford Work/WELL/Data/Data_US/Recoded/"
#setwd(my_depo)
# DF_US %>% saveRDS("2021_0115_WELL_US_n_6245_recoded.rds")

#### Project - Wearable Device ####
my_depo = "C:/Users/WeiKuang/Desktop/Stanford Work/WELL/Data/Data_US/Recoded/"
setwd(my_depo)
DF_US %>% saveRDS("2020_0226_WELL_US_n_6245_recoded.rds")
```


#### Variables related to NAFLD ####

liver_upper
liver_anterior
liver_oblique
liver_surface
liver_edge
liver_parenchyma
liver_vessel
liver_envelope
fatty_liver_b
fatty_liver
pic1
pic2
liver_data_recovered
liver_data_missing



